Deep mutational scanning of an H3 hemagglutinin can inform evolutionary forecasting of human H3N2 influenza virus

### 2017-12-28 paper/manuscript.tex
Switched order of distributions in distance_distribution fig
Add G78D-T212I supplemental figure

\subsection*{Deep mutational scanning of HA from a recent strain of human H3N2 influenza virus}
We performed a deep mutational scan to measure the effects of all amino-acid mutations to the A/Perth/16/2009 (H3N2) HA on viral replication in cell culture. 
This strain was the H3N2 component of the influenza vaccine from 2010-2012.
We used a variant of this HA that supported better viral replication in cell culture than did the sequence in GenBank \comment{Figure S1}.
This HA variant contained two mutations: G78D and T212I.
The G78D mutation occurs at low frequency in natural H3N2 sequences, and T212 is a site where a mutation to Ala rose to fixation in human influenza in $\sim$2011.
%We first generated virus carrying the wildtype Perth/2009 HA and NA and the WSN internal genes by reverse genetics, and serially passaged the virus in cell culture a total of six times.
%This passaging was completed to allow the virus to adapt to growth in cell culture and to find cell-culture adaptation mutations that emerged during the course of passaging.

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/trunkvssidebranch/trunkvssidebranch.pdf}}
\caption{\label{fig:trunkvssidebranch}
{\bf Mutations in evolutionary successful strains tend to be more favorable than in strains that die out.}
(A) Phylogenetic tree of human H3N2 influenza virus from 1968-present. 
The trunk is shown in red, and side branches are shown in blue.
The gray branches represent the part of the tree for which we cannot yet distinguish the trunk from side branches.
(B) Using the Perth/2009 H3 preferences, we calculated the log$_{2}$ mutational effect for trunk and side branch mutations in windows of 5 years for every year from 1968-2013. 
The median log$_{2}$ mutational effect in a given window is shown as circles for trunk mutations and triangles for side branch mutations. 
The shaded region demarcates the interquartile range of trunk and side branch mutational effects.
The median trunk mutational effects are consistently higher than the median side branch mutational effects for all windows.
(C) The log$_{2}$ mutational effect for all side branch and all trunk mutations (left panel), in addition to all mutations in internal nodes and terminal nodes on the side branches (right panel) are shown.
The preferences were randomized for a number of iterations to estimate significance.
The effects of trunk mutations are higher than side branch, internal and terminal node, mutations.
}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/sequence_preference/sequence_preference.pdf}}
\caption{\label{fig:sequence_preference}
{\bf The trunk exhibits higher sequence preference than side branches.}
figure text
}
\end{figure}

@@

>>>>

\subsection*{Deep mutational scanning of HA from a recent strain of human H3N2 influenza virus}
We performed a deep mutational scan to measure the effects of all amino-acid mutations to the A/Perth/16/2009 (H3N2) HA on viral replication in cell culture. 
This strain was the H3N2 component of the influenza vaccine from 2010-2012.
We used a variant of this HA that supported better viral replication in cell culture than did the sequence in GenBank (Figure~\ref{suppfig:Perth2009_mut}).
This HA variant contained two mutations: G78D and T212I.
The G78D mutation occurs at low frequency in natural H3N2 sequences, and T212 is a site where a mutation to Ala rose to fixation in human influenza in $\sim$2011.
%We first generated virus carrying the wildtype Perth/2009 HA and NA and the WSN internal genes by reverse genetics, and serially passaged the virus in cell culture a total of six times.
%This passaging was completed to allow the virus to adapt to growth in cell culture and to find cell-culture adaptation mutations that emerged during the course of passaging.

@@

\begin{figure}
\centerline{\includegraphics[width=0.75\textwidth]{figs/trunkvssidebranch/trunkvssidebranch.pdf}}
\caption{\label{fig:trunkvssidebranch}
{\bf Mutations in evolutionary successful strains tend to be more favorable than in strains that die out.}
(A) Phylogenetic tree of human H3N2 influenza virus from 1968-present. 
The trunk is shown in red, and side branches are shown in blue.
The gray branches represent the part of the tree for which we cannot yet distinguish the trunk from side branches.
(B) Using the Perth/2009 H3 preferences, we calculated the log$_{2}$ mutational effect for trunk and side branch mutations in windows of 5 years for every year from 1968-2013. 
The median log$_{2}$ mutational effect in a given window is shown as circles for trunk mutations and triangles for side branch mutations. 
The shaded region demarcates the interquartile range of trunk and side branch mutational effects.
The median trunk mutational effects are consistently higher than the median side branch mutational effects for all windows.
(C) The log$_{2}$ mutational effect for all side branch and all trunk mutations (left panel), in addition to all mutations in internal nodes and terminal nodes on the side branches (right panel) are shown.
The preferences were randomized for a number of iterations to estimate significance.
The effects of trunk mutations are higher than side branch, internal and terminal node, mutations.
}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=0.75\textwidth]{figs/sequence_preference/sequence_preference.pdf}}
\caption{\label{fig:sequence_preference}
{\bf The trunk exhibits higher sequence preference than side branches.}
figure text
}
\end{figure}

@@

\section*{Supplementary Material}
\FloatBarrier
\pagenumbering{arabic}% resets `page` counter to 1
\renewcommand*{\thepage}{S\arabic{page}}

\begin{suppfigure}
\centerline{\includegraphics[width=\textwidth]{figs/S01_G78D-T212I/G78D-T212I.pdf}}
\caption{\label{suppfig:Perth2009_mut}
{\bf Characterization of the G78D-T212I Perth/2009 HA variant.} 
(A) 
The G78D-T212I Perth/2009 HA variant grows to higher titers than do viruses carrying the wildtype Perth/2009 HA.
Each virus was generated in duplicate by reverse genetics and passaged once at MOI = 0.01 in MDCK-SIAT1-TMPRSS2 cells.
The rescue and passage viral supernatants were titered, with the points marking each duplicate and the bar marking the mean.
(B)
The D78 variant remained at a low frequency in natural human H3N2 sequences over the past $~\sim$10 years.
The A212 variant rose to fixation in $~\sim$2011, replacing the T212 variant.
}
\end{suppfigure}
